---
title: "Read Gallery"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Read Gallery}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = T, comment = "#>")
options(tibble.print_min = 4L, tibble.print_max = 4L)
library(dplyr)
library(readGallery)
library(readr)
library(ggplot2)
```

## Introduction

This `readGallery` package allows you to read *art* data files with the *gallery* library (see http://art.fnal.gov). 

### System requirements

The system requirements are as follows:

* You need a working gallery environment including the accompanying Root and Python packages. Be sure this environment is set up and active when you start *R* or *RStudio*. 
* If you are getting data from Fermilab dCache, you will need the appropriate proxy
* *R* and *RStudio* will pick up `$PATH` and `$LD_LIBRARY_PATH` but not other necessary environment variables. You need to write an `~/.REnviron` file set the variables within *R*. You can write this file from the bash command line (do this before starting *R* or *RStudio*) with,

```bash
$ env | egrep '(_INC|_LIB=|_DIR=|^MRB=|^PYTHON)' > ~/.Renviron
```

* You also need the `reticulate` *R* package. See https://github.com/rstudio/reticulate for information and installation instructions.

### Workflow

The basic workflow you need to follow is:

1. Load the `readGallery` library.
2. Locate your data files and put the paths in a vector. If you are using `XRootD`, then to get the appropriate URI(s) you should look-up the file paths with `ifdh ls` and then run them through `readGallery::xrootify`. 
3. Declare use of an *art* data product with `readGallery::useDataProduct`. The argument is a string with the C++ class name including namespace if necessary. Be sure to wrap in `std::vector< >` if necessary.
4. Create the *art* input tags you need with `readGallery::artInputTag`. You can determine the input tags of objects in the data file with the *art* binary `product_sizes_dumper`.
5. Write a python class to serve as the *reader*. See below. 
6. Run `readGallery::getGalleryData` to process the files.
7. Run `readGallery::galleryReader_df` on the reader object(s) to extract the collected data as an R data frame.

That all looks like a lot, but it's pretty easy. Let's try it!

## Locate  data files

I happen to have some data files in Fermilab dCache (note that you'll need the proper environment and proxy established).

```{r}
system('ifdh ls /pnfs/GM2/scratch/users/lyon/arr_20170307/*/*.root | grep .root | grep _10k', intern=T) %>% 
  xrootify() -> myFiles
myFiles
```

Let's do one file for testing
```{r}
myOneFile <- myFiles[1]
myOneFile
```

## Declare use of a data product

We are going to read the `GhostDetectorArtRecord` objects. We need to declare use of this object.

```{r}
useDataProduct('std::vector<gm2truth::GhostDetectorArtRecord>')
```

We would do the same for other objects too if needed.

## Create input tags

There are two instances of `GhostDetectorArtRecord` objects in the file. We'll load both (eventually). Here are their input tags.  
```{r}
gh_cyl_tag <- artInputTag('artg4:GhostCylinderDetector')
gh_nwd_tag <- artInputTag('artg4:GhostNearWorldDetector')
```

## Write the reader class

You write a reader class in **Python**. The reader class should satisfy the following.

* It must have `prepare` and `fill` methods. The `prepare` method gets the reader ready before file processing, perhaps by getting the `get_valid_handle` function from the `gallery.Event`. The `fill` method does everything necessary to fill the value rows with data from the `gallery.Event`. 
* It should have `values` (the accumulated data from the files) and `colnames` (names of the columns) methods to return those data. If you do not provide those methods, then you will need to use `readGallery::galleryReader_df` with the `values` and `colnames` arguments and provide methods to extract that data. 

You can use a base class, `galleryReader.GalleryReaderBase` to do some of the boilerplate things for you. 

You can either write the python class in a file and import it with `readGallery::createReaderClass_from_file` or write it in an *R* string and import it with `readGallery::createReaderClass_from_string`. Both of these functions return the main python environment. You then need to extract your class from the environment with the `$<className>` operation. For example,

```{r eval=FALSE}
createReaderClass_from_file('myReaderClass.py')$MyReader   # class MyReader  is in the python file
```

There is one example Reader class in this package for `GhostDetectorArtRecord`. It uses the base class. Here is the base class for reference.

```python
class GalleryReaderBase:
  """A base class for simple readers"""

  def __init__(self, inputTag):
    self.vals = []
    self.inputTag = inputTag
    self.getValidHandle = None   # Shouild be set in the prepare method
    self.names = None # Needs to be set in derived class; self.names = [...]

  def colnames(self):
    return self.names

  def values(self):
    return self.vals

  def prepare(self, ROOT, ev):
    self.vals = []  # Protect against re-run
    # Your code sets self.getValidHandle

  def fill(self, ROOT, ev):
    # Your code fills self.vals
    return True
```
We can make a skeleton class and fill in the details. Most of the time the skeleton will work fine, but you will want to make changes. Perhaps remove some columns. Add constraints. Other things. Remember that the philosophy is to load only the data you **need**. 
```{r}
readerClassSkel('gm2truth::GhostDetectorArtRecord', writeFile = 'ghostDetectorRecord.py')
```


We can display the `GhostDetectorArtRecordReader` class with the `readr::read_file(...) %>% cat` pipeline.  

```{r}
readr::read_file( 'ghostDetectorRecord.py') %>% cat
```

Let's load it into python. Note the use of `$<class name>`,
```{r}
createReaderClass_from_file('ghostDetectorRecord.py')$GhostDetectorArtRecordReader -> GHReader
```

Make the reader objects
```{r}
ghCReader <- GHReader(gh_cyl_tag)
```

## Process the files

We are now ready to process the files! Let's just do one at first. 
```{r}
getGalleryData(myOneFile, ghCReader)
```

```{r}
ghcdf <- galleryReader_df(ghCReader) %>% tbl_df
ghcdf
```
Let's try two readers (we'll use the same Reader class, but initialize it with a different input tag).
```{r}
ghNReader <- GHReader(gh_nwd_tag)
```

Let's read in the data again
```{r}
getGalleryData(myOneFile, c(ghCReader, ghNReader))
```

```{r}
ghndf <- galleryReader_df(ghNReader) %>% tbl_df
ghndf
```

Let's try all the data! This will take longer. 
```{r}
getGalleryData(myFiles, c(ghCReader, ghNReader) )
```

And look at the data
```{r}
ghcdf <- galleryReader_df(ghCReader)
ghndf <- galleryReader_df(ghNReader)
```

How many rows did we get per file?
```{r}
ghcdf %>% group_by(fileEntry) %>% tally()
```
```{r}
ghndf %>% group_by(fileEntry) %>% tally()
```
## Timing information

The function `readGallery:getGalleryData` can return timing information in the form of a python object. For example,

```{r}
times <- getGalleryData(myOneFile, ghCReader)
```

```{r}
times$allTime
```
Look at time it took to process events. 
```{r}
et <- times$eventTimes
mean(et)
```
```{r}
qplot(seq_along(et), et) + xlab('eventEntry') + ylab('Processing time (s)')
```

The slow ones seem to be in the beginning
```{r}
ets <- et[100:length(et)]
qplot(seq_along(ets), ets) + xlab('eventEntry') + ylab('Processing time (s)')
```




